---
output: rmarkdown::github_document
---

deepregression
======
[![Build Status](https://travis-ci.com/davidruegamer/deepregression.svg?token=UQrq5mDFv9MYhofnJ2xn&branch=master)](https://travis-ci.com/davidruegamer/deepregression)


```{r setup, include=FALSE, cache=TRUE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

# Installation

Since the repository is still private, clone the repository to your local machine
and run the following

```{r, include=FALSE, message=FALSE}
library(devtools)
load_all("R")
```

Also make sure you have installed all the dependencies:

* Matrix
* dplyr
* keras
* mgcv
* reticulate
* tensorflow
* tfprobability

In the future, the package can be installed as follows:

To install the latest version of deepregression:

```{r, eval=FALSE}
library(devtools)
install_github("davidruegamer/deepregression")
```

# How to cite this?

Until published, please cite the following preprint:

```
@article{rugamer2020unifying,
  title={A Unifying Network Architecture for Semi-Structured Deep Distributional Learning},
  author={R{\"u}gamer, David and Kolb, Chris and Klein, Nadja},
  journal={arXiv preprint arXiv:2002.05777},
  year={2020}
}
```

# How to use this?

See [the tutorial](vignettes/tutorial.md) for a detailed introduction.

